Generality in food webs scales with species richness, not latitude

### 2014-07-09 manuscript/latitudepaper.tex
Roughed out discussion, finished results. Still need to do figures.

@@

\subsection*{Effect of Latitude}

@@

Once the best formulation of the relationship between each property and $S$ had been determined, we then assessed
the impact of latitude on this relationship. There are two ways that properties might be affected by latitude:
1) the slope of the scaling relationship with $S$ might vary over latitude or 2) the residuals $\epsilon$ might
vary over latitude. We were primarily interested in the first possibility, and so included a term for the effect of 
latitude on the scaling exponent $\beta$ in equation \ref{Power}. In addition, as other studies have suggested that 
food webs in different ecosystem types have characteristic structural differences, we included a dummy variable for 
each ecosystem type (stream, lake, terrestrial, or marine, with estuary used as the `intercept' ecosystem type). 
We also considered other trends in the data that could affect relationships between latitude and species richness.
To account for changing standards in data collection over time \citep{Dunne2006} we included an effect of year of
publication. As 11 of our XX webs included humans, and therefore might have different properties than other networks,
we included a factorial effect of the presence of humans. Finally, some of our webs describe the same area in different
seasons or different years. To account for likely correlation among these webs, we included a random effect of site ID.

@@

\begin{equation}
\label{PowerLat}
$Z_{i}=\alpha S^{\beta_{0}+\beta_{1}\delta_{i}+\beta_{2}\Ecotype+\beta_{X}\delta_{i}\Ecotype}\Year_{i}\Humans_{i} + \epsilon_{i} $ \,,

@@

where $\delta$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction), 
$\beta_{1}$ represents the effect of latitude on the scaling exponent $\beta$ in equation \ref{Power}. If the scaling
relationship with S for a property suggested that lognormal error was more appropriate, we used the logarithmic re-
expression:

\begin{equation}
\label{LogLat}
$Z_{i}= log(\alpha)+\beta_{0}log(S) + \beta_{1}log(S)\delta_{i} +\beta_{2}log(S)Ecotype + \beta_{X}log(S)\delta_{i}Ecotype +log(year_{i}) +log(Humans_{i})  +\epsilon_{i} $ \,.

We calculated the AIC of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models.
Simplified models were obtained by systematically removing all possible combinations of terms from the full model.
The term for species richness was not removed from a model if any of the terms affecting the exponent $\beta$ were 
included in the model. The model with the lowest AIC was selected as the best-fitting model. For power-law formulations
this simplification was performed by hand. For logarithmic formulations, this simplification was performed using the
R \citep{R} function dredge from package MuMIn \citep{MuMIn}. We then used the R \citep{R} function lmer
from the package lmerTest \citep{lmerTest} to estimate the standardized effects ($\beta$s) for each fixed effect in the 
best-fitting models, as well as their corresponding $p$-values.

@@

@@

In this data set, none of the best fit models for species richness, connectance, or links per species included an 
effect of latitude, although all three properties varied across ecosystem types (Table 2).

The best-fit models for most food-web properties exhibited scaling relationships with species-richness and/or 
connectance. Notable exceptions were the proportions of basal resources, herbivores, and omnivores, which varied only 
across year published; and the proportions of top predators and intermediate consumers, which varied with connectance 
and species richness and connectance (respectively) although not significantly.

For the food-web properties that did scale with species-richness or connectance, most of these relationships were 
constant over latitude. Latitude terms were not included in the best-fit models for SD of generality or 
SD of links. Best-fit models for vulnerability, SD of vulnerability, and path length included latitude terms but showed 
very weak relationships ($\beta$=3.52 x $10^{-17}$, $p$=0.170; $\beta$=-0.001, $p$=0.050; and $\beta$=0.005, $p$=0.069; 
respectively). None of the above models included an effect of latitude on scaling with S or C.

The scaling rules of mean trophic level and generality, however, did vary over latitude. 
Mean trophic level increased with latitude ($\beta$=0.011, $p$<0.001) and with species richness 
($\beta$=0.633, $p$<0.001), but both trends were weakened by an interaction between latitude and species-richness 
($\beta$=-0.008, $p$=0.001). The best-fitting model for mean trophic 
level did not include a connectance term. The increase in mean trophic level with species richness was highest for 
estuarine food webs ($\beta$=0.633, $p$<0.001) and lower for freshwater ($\beta$=0.259, $p$=0.002), terrestrial 
($\beta$=0.234, $p$=0.005), and marine food webs ($\beta$=0.152, $p$<0.001). For food webs with high species richness, 
trophic level declined with latitude while trophic level in species-poor webs showed the opposite trend 
(Fig. \ref{TL_all}). The number of species for which the slope of trophic level over latitude changed from negative to 
positive varied across ecosystem types.

Generality showed the opposite trend, increasing with latitude for webs with high species richness and decreasing with
latitude for species-poor webs (Fig. \ref{gen_all}).
This increase was driven by a positive interaction between latitude and species richness ($\beta$=0.014, $p$=0.002), as 
the main effect of latitude was negative ($\beta$=-0.018, $p$=0.002). Generality also decreased with species richness 
for estuarine food webs ($\beta$=-0.150, $p$=0.002) but increased with species richness for freshwater, terrestrial, 
and marine webs ($\beta$=0.453, $p$=0.005; $\beta$=0.523, $p$=0.006; and $\beta$=0.675, $p$=0.001; respectively). 
Generality increased with connectance equally in all ecotypes and across all latitudes ($\beta$=0.799, $p$<0.001).

@@

The tendency of food-web structure to exhibit scaling relationships with both species richness and connectance has been 
well-established \citep{,,,,}. As species richness in particular is also known to vary systematically over latitude 
\citep{,,,,}, intuitively one might suspect that the scaling relationships of other food web properties might also vary
over latitude \citep{}. However, we found that only five of 17 food web properties varied significantly over latitude, 
and that only, of these, only mean and maximum trophi level and generality showed different scaling laws at different latitudes.

The lack of latitudinal gradients in most of the food-web properties considered here may be due to the lack of a 
latitudinal gradient in species richness in this data set. If so, that strongly implies that the latitudinal gradients 
that were observed were driven by some other factor than latitudinal changes in species richness. As we observed 
increasing generality away from 
the equator in conjunction with decreasing trophic level (at least in high-S food webs), it appears that species nearer
the top of the food web may be feeding on a wider variety of prey at the poles, lowering their trophic levels. 
As both feeding on many prey within a trophic level (non-omnivorous generality) \citep{Romanuk2006} and feeding on many 
trophic levels \citep{} have been shown to decrease population variability, high generality at higher latitudes may act 
as ``insurance'' for upper-level consumers by allowing them to switch prey according to availability.

This contrasts with an earlier study of plant-animal mutualistic networks which found that generality increased in the 
tropics \citep{Schleuning2012}. That increase in generality was attributed to increasing species richness, as high 
resource diversity might lead to generalized consumers \citep{Schleuning2012}. While we also found increasing 
generality with increasing species richness in our dataset, it seems unlikely that higher species richness at the poles 
is driving the observed trend in generality. This suggests that while mutualistic and antagonistic interactions may
be similarly affected by species richness, they respond differently to other latitudinally-associated variables.

[[High TL/low generality in species-poor food webs?]]

% A contrary argument was proposed in an earlier study \citep{Vazquez2004}. In this study, it was noted that bipartite 
% networks such as plant-pollinator and host-parasite networks are typically highly nested (specialists interact with 
% subsets of the interaction partners of generalists), and that nestedness tends to increase with network size 
% \citep{Vazquez2004}. Therefore, as the number of species in a community increases the number of extreme specialists in 
% a community should increase and generality should decrease. Importantly, this argument suggests that latitude can 
% affect generality only indirectly via species richnes. As the networks in our dataset did not increase in size towards 
% the tropics, and there was an effect of latitude as well as species richness, it seems unlikely that either of these 
% two arguments holds.

@@

While the overall consistency of scaling laws between species richness and connectance and other food-web properties is 
remarkable, there are nonetheless signification variations in the scaling laws of trophic level and generality. These food web 
properties vary both over latitude and across different network types. [[something punchy]]

>>>>

\subsection*{Relationships with Latitude}

We first examined simple linear relationships between latitude and $S$, $Z$, $G$, and $V$. We fit models of the form

\begin{equation}

S_{i} = \alpha_{0} + \alpha_{1}\Delta_{i}+\epsilon_{i} \,,

where $S_{i}$ is the species richness of web $i$, $\Delta_{i}$ its absolute latitude (degrees north or south 
regardless of direction), and $\epsilon_{i}$ a residual error term.
These simple linear models, fit using the function lm in R \citep{R} package stats, were used to confirm that there are indeed latitudinal gradients in food-web structure in this data set.

@@

\subsection*{Effect of Latitude on Scaling}

@@

Once the best formulation of the relationship between each property and $S$ had been determined, we then assessed
the impact of latitude on this relationship. There are two ways that properties might be affected by latitude:
1) the slope of the scaling relationship with $S$ might vary over latitude or 2) the residuals $\epsilon$ might
vary over latitude. To account for both possibilities we included terms for the effect of latitude on the scaling
exponent $\beta$ in equation \ref{Power} as well as on the residual. In addition, as other studies have suggested that 
food webs in different ecosystem types have characteristic structural differences, we included a dummy variable for 
each ecosystem type (stream, lake, terrestrial, or marine, with estuary used as the `intercept' ecosystem type). 
We also considered other trends in the data that could affect relationships between latitude and species richness.
To account for changing standards in data collection over time \citep{Dunne2006} we included an effect of year of
publication. As 11 of our XX webs included humans, and therefore might have different properties than other networks,
we included a factorial effect of the presence of humans. Finally, some of our webs describe the same area in different
seasons or different years. To account for likely correlation among these webs, we included a random effect of site ID.

@@

\begin{equation}
\label{PowerLat}
$Z_{i}=\alpha S^{\beta_{0}+\beta_{1}\Delta_{i}+\beta_{3}\Sigma_{i} + \beta_{4}\Lambda_{i} + \beta_{5}\Tau_{i} + \beta_{6}\Mu_{i} 
+ \beta_{7}\Delta_{i}\Sigma_{i} + \beta_{8}\Delta_{i}\Lambda_{i} + \beta_{9}\Delta_{i}\Tau_{i} + \beta_{10}\Delta_{i}\Mu_{i}}
+ \rho{1}\Delta_{i} + \rho_{2}\Upsilon_{i} + \rho_{3}\Eta_{i} + P_{i} + \epsilon_{i} $ \,,
\end{equation}

where $\delta$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
$\Sigma_{i}$, $\Lambda_{i}$, $\Tau_{i}$, and $\Mu_{i}$ are 
factors indicating whether web $i$ is a stream, lake, 
terrestrial, or marine food web (e.g., $\Sigma_{i}$=1 for 
a stream food web, 0 otherwise). If $\Sigma_{i}$, $\Lambda_{i}$, $\Tau_{i}$, and $\Mu_{i}$ are all 0, web $i$ 
describes an estuary food web. $\Upsilon_{i}$ is the year web $i$ was published, $\Eta_{i}=1$ if web $i$ contains humans, 0 otherwise, and $P_{i}$ is a random effect of 
site ID. Note that all fixed effects $\beta$ refer to factors acting on the scaling exponent between $Z$ and $S$
while fixed effects $\rho$ refer to factors acting on the residual error.

The logarithmic formulation of this model is:

\begin{equation}
\label{LogLat}
$Z_{i}= log(\alpha)+\beta_{0}log(S) + \beta_{1}\Delta_{i}log(S) +\beta_{3}\Sigma_{i}log(S) + \beta_{4}\Lambda_{i}log(S)  + \beta_{5}\Tau_{i}log(S)  + \beta_{6}\Mu_{i}log(S)  
+ \beta_{7}\Delta_{i}\Sigma_{i}log(S)  + \beta_{8}\Delta_{i}\Lambda_{i}log(S)  + \beta_{9}\Delta_{i}\Tau_{i}log(S)  + \beta_{10}\Delta_{i}\Mu_{i}log(S) 
+ \rho_{1}\Delta_{i} + \rho_{2}\Upsilon_{i}) + \rho_{3}\Eta_{i} +\epsilon_{i} $ \,.

@@

As in the power-law formulation, fixed effects $\beta$ 
affect the scaling exponent and fixed effects $\rho$ 
affect the (lognormal) residual.

We calculated the AIC of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models.
Simplified models were obtained by systematically removing all possible combinations of terms from the full model.
The term for species richness was not removed from a model if any of the terms affecting the exponent $\beta$ were 
included in the model. The model with the lowest AIC was selected as the best-fitting model. For power-law formulations
this simplification was performed by hand. For logarithmic formulations, this simplification was performed using the
R \citep{R} function dredge from package MuMIn \citep{MuMIn}. We then used the R \citep{R} function lmer
from the package lmerTest \citep{lmerTest} to estimate the standardized effects ($\beta$s and $\rho$s) for each fixed effect in the 
best-fitting models, as well as their corresponding $p$-values.

@@

\subsection*{Relationships with Latitude}

$S$ and $G$ both increased with increasing latitude, contrary to expected latitudinal gradients (Table \ref{Latlms}). $G$ and $V$ were not significantly correlated with latitude.

\label{Latlms}
\caption{Intercepts and slopes for simple linear regressions of food web properties against absolute latitude.}
\begin{table}{l | l l l l}
\multirow{2}Property & \multicolumn{2}{c}{Intercept} & \multicolumn{2}{c}{Slope} \\
 & $\alpha$ & $p$-value & $\alpha$ & $p$-value \\
\hline
$S$ & 23.10 & \textless0.001 & 0.284 & 0.040 \\
$Z$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
$G$ & 3.269 & \textless0.001 & 0.049 & 0.014 \\
$V$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
\end{table}
\end{tabular}

@@

In this data set, none of the best fit models for species richness, connectance, or links per species included an 
effect of latitude on either the scaling exponent or residual (Table \ref{bestfits}). 
$G$ increased slightly in more recent webs ($\rho{2}$=0.005, $p$<0.001), but the best-fit models for $V$ and $Z$ did not include a term for year of publication. 
In contrast, $Z$ and $V$ were higher in lake food webs ($\beta_{4}$=0.10, $p$<0.001 and $\beta_{4}$=0.10, $p$<0.001, respectively) while the best-fit model for $G$ did not include an effect for any ecosystem type.

@@

The tendency of food-web structure to exhibit scaling relationships with both
species richness and connectance has been  well-established \citep{,,,,}. As
species richness in particular is also known to vary systematically over
latitude  \citep{,,,,}, intuitively one might suspect that any relationship
between food-web properties such as generality might be due to the latitudinal
gradient in species richness. Here, we found that while there were latitudinal
gradients in species richness and generality, latitude was not related to
either the scaling relationship between generality and species richness or the
residuals of this scaling relationship.

As the latitudinal gradient in species richness in this data set ran in the
opposite direction to the general trend for angiosperms \citep{}, mammals
\citep{}, birds \citep{}, marine organisms? \citep{}, we conclude that it is
likely an artifact of the data set. Academic institutions have the longest
history in temperate latitudes, and it may be that communities at these
latitudes are better understood or more completely sampled. We do not believe
that our observation of higher species richness at higher latitudes represents
a true biological trend. As such, it does not seem likely that the increase in
generality with latitude that we observed is due to an underlying tendency for
species to pursue a wider variety of prey at higher latitudes. Rather, the
trend in generality can be more parsimoniously explained by the trend in
species richness.

Our findings here add further doubt to the niche-breadth hypothesis, as
narrower niches in the tropics were associated with lower species richness in
this data set. It appears instead that, as more species are available, each
species tends to consume more prey, be consumed by more predators, and have
more links in general (check with figs). This makes intuitive sense as a more
species-rich community is likely to have more species with similar behaviour
and morphology that thereby have overlapping predator or prey sets.

Parallel increases in generality and species richness across latitude have
also been noted in studies of plant-pollinator networks (although in that case
species richness was highest in the tropics) \citep{Schleuning2012}. Although
the authors of that study do not explicitly examine the relationship between
specialization and network size, they do cite increasing plant species
richness as a potential driver for increasing generalization
\citep{Schleuning2012}. Although the scaling of mutualist network structure
does not appear to have been explored as systematically as that of unipartite
food webs, some poperties such as nestedness \citep{Vazquez2004} and XX
\citep{} have been shown to scale with species richness. As such scaling
appears to be common to both antagonistic and mutualistic food webs, and
trends in species richness can explain counterintuitive trends in network
structure \citep{Schleuning2012}, it seems likely that many latitudinal trends
in network structure are at least partly due to trends in species richness.

%Nestedness increases with A+P, increases number of extreme specialists AND number of generalists. Not clear how degree should scale with S according to scaling with nestedness.

@@

It is not our intention here to question the validity of the wide range of
studies exploring variation in network structure across changing environments.
However, if those studies have not taken into account changes in species
richness over environmental gradients, it is possible that observed changes in
food-web structure are due only indirectly to the environmental variables of
interest. Effects of environmental variables on network structure should
therefore be interpreted with caution unless effects of changing species
richness are made explicit.
